---
title: "PA1_template"
author: "Gustavo García"
date: "10/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data
```{r load}
#Name file downloaded
name <- "activity.zip"

# Check if file already exists
if (!file.exists(name)){
  URL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
  download.file(URL, name, method="curl")
} 

# Check if folder exists
if (!file.exists("activity.csv")) { 
  unzip(name) 
}
```

## Reading the data
```{r read}
activity<- read.csv("activity.csv")
```


## 1. Preprocessing the data
```{r preprocessing}
library(ggplot2)
library(lubridate)
activity<- transform(activity, date= as.Date(date, format="%Y-%m-%d"))
```

## 2. Histogram of the total number of steps taken each day

```{r plot1}
p1<- ggplot(data=activity, aes(x=date, y=steps)) 
p1 + geom_bar(stat="identity", fill="steelblue")
```

## 3. Mean and median number of steps taken each day

```{r plot2}
aggregate(steps~date, activity, FUN = function(x) c(mean = mean(x), median = median(x)))
```

## 4. Time series plot of the average number of steps taken
```{r plot3}
df<- aggregate(steps~interval, activity,mean)
p2<- ggplot(data=df, aes(x=interval, y=steps)) 
p2 + geom_line()
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
df[which.max(df$steps),]
```

## 6. Code to describe and show a strategy for imputing missing data

1.  Total number of missing values in the dataset
```{r}
sum(!complete.cases(activity))
```

2. Estrategy

```{r}
NASteps<- which(is.na(activity$steps))
activity[NASteps,"steps"]<- sapply(activity[NASteps,"interval"],                                    function(x) (df[which(df$interval==x),"steps"]))
```

## 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r plot4}
p3<- ggplot(data=activity, aes(x=date, y=steps)) 
p3 + geom_bar(stat="identity", fill="steelblue")
```

Mean and median number of steps taken each day

```{r}
aggregate(steps~date, activity, FUN = function(x) c(mean = mean(x), median = median(x)))
```
## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r plot5}
weektype <- function(date){
  if (weekdays(date) %in% c("Saturday","sunday","sabado","domingo"))
    type<- "weekend"
  else{
    type<- "weekday"
  }
   type
}
activity$type<- factor(sapply(activity$date, weektype))
df2<- aggregate(steps~interval+type, activity,mean)
p2<- ggplot(data=df2, aes(x=interval, y= steps)) 
p2 + facet_wrap(type~., nrow=2, ncol=1)+ geom_line()
```
